Supplementary Information for Diverse lithic production strategies in southwest China during Late Middle Pleistocene

### 2020-10-04 analysis/paper/paper.Rmd
nice looking table 1

```{r}
point_alpha <- 0.1
```

@@

# clean data to help with analysis
flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")
flakes <- flakes[!flakes$material %in% c('sandstone', 'quartz'),]

@@

@@

```{r}
# is it significant?
t_test_str_l_mass <- 
apa_print(t.test(data = flakes_L, mass ~ L))$statistic %>% 
  str_remove_all("\\$|\\{|\\}|\\\\|mathit")

>>>>

@@

# clean data to help with analysis
flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")
flakes <-
  flakes[!flakes$material %in% c('sandstone', 'quartz'), ] %>%
  mutate(complet_flk = ifelse(
    type1 %in% c(
      "flake",
      "blade",
      "KBW",
      "tablet",
      "leva",
      "leva?",
      "DBD",
      "crest"
    ),
    "complete flake",
    ifelse(
      type1 %in% c("flake brk", "blade brk", "LVFB", "proximal"),
      "flake breaks",
      "other"
    )
  ))

@@

```{r}
flakes_complete_non_Lev <- 
  flakes_L %>% # only complete flakes for non-Levallois
  filter(complet_flk == "complete flake")

flakes_Lev  <- 
  flakes_L %>% 
  filter(L == "Levallois")

# combine both here
flakes_Lev_non_Lev <- 
bind_rows(flakes_complete_non_Lev,
          flakes_Lev) %>% 
  as_tibble()

# check the CVs
CV <- function(the_vector){
  mean_ <- mean(the_vector, na.rm = TRUE)
  sd_ <- sd(the_vector, na.rm = TRUE)
  cv <- (sd_/mean_) * 100
  round(cv, 3)
}

# prepare mean and sd functions
mean_t <- function(x) mean(x, na.rm = TRUE)
sd_t <- function(x) sd(x, na.rm = TRUE)

# for the table 'Results of descriptive statistics for Levallois and non-Levallois flakes. ‘PLF’ stands for preferential Levallois flake; ‘CF’ stands for complete flake.'
compare_lev_non_lev_vars_tbl <- 
flakes_Lev_non_Lev %>% 
  as_tibble() %>% 
  select(L, where(is.numeric)) %>% 
  group_by(L) %>% 
  summarise(across(where(is.numeric),
            list(mean = mean_t,
                 sd = sd_t,
                 cv = CV),
            .names = "{.fn}.{.col}")) %>%
  pivot_longer(-L) %>%
  separate(col = name,
           into = c("stat", "var"),
           sep = "\\.") %>% 
  mutate(L = ifelse(L == "Levallois",
                    "PLF",
                    "CF")) %>% 
  unite("lev_stat", c(stat, L), 
        sep = "_") %>% 
  pivot_wider(names_from = lev_stat ) %>% 
  mutate(difference_in_cv = cv_PLF - `cv_CF`) %>% 
  mutate(across(where(is.numeric), round, 2))

library(gt) 
tab <- 
  compare_lev_non_lev_vars_tbl %>%
  gt(rowname_col = "var") %>% 
  tab_spanner(
    label = "Mean (mm)",
    columns = vars(mean_PLF, 
                   mean_CF)) %>% 
  tab_spanner(
    label = "SD",
    columns = vars(sd_PLF, 
                   sd_CF)  
  ) %>% 
  tab_spanner(
    label = "CV (%)",
    columns = vars(cv_PLF, 
                   cv_CF,
                   difference_in_cv) 
  )

Are CV values lower for Levallois flakes than for non-Levallois flakes?

```{r}
compare_lev_non_lev_vars_tbl  %>% 
  select(contains("cv_")) %>% 
  stack() %>% 
  wilcox.test(values ~ ind, 
              data = .) %>% 
  apa_print() %>% 
  .$statistic %>% 
  str_remove_all("\\$|\\{|\\}|\\\\|mathit")

@@

```{r}
# plotting setting
point_alpha <- 0.1

# is it significant?
t_test_str_l_mass <- 
apa_print(t.test(data = flakes_L, mass ~ L))$statistic %>% 
  str_remove_all("\\$|\\{|\\}|\\\\|mathit")
